### 2021-03-23 README.Rmd
First creation of the mSHAP function, some documentation and other package setup

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# mshap

<!-- badges: start -->
<!-- badges: end -->

The goal of mshap is to allow SHAP values for two-part models to be easily computed.